---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonmem2rx

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nonmem2rx/branch/main/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nonmem2rx?branch=main)
[![CRAN version](http://www.r-pkg.org/badges/version/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CodeFactor](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx/badge)](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx)
![r-universe](https://nlmixr2.r-universe.dev/badges/nonmem2rx)
<!-- badges: end -->

The goal of nonmem2rx is to convert a NONMEM control stream to
`rxode2` for easy clinical trial simulation in R.

## Installation

You can install the development version of nonmem2rx from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nlmixr2/nonmem2rx")
```

## Example

Once `nonmem2rx` has been loaded, you simply type the location of the
nonmem control stream for the parser to start.  For example:

```{r example}
library(nonmem2rx)
mod <- nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res", save=FALSE)
mod
```

The process steps are below:

- Read in the nonmem control stream and convert the model to a
  `rxode2` ui function.

- Try to determine an endpoint in the model (if possible), and convert
  to a fully qualified ui model that can be used in `nlmixr2` and
  `rxode2`.

- If available, `nonmem2rx` will read the final parameter estimates
  and update the model. (See below for source)

- This will read in the nonmem input dataset, and search for the
  output files with `IPRED`, `PRED` and the `ETA` values. The
  translated `rxode2` model is run for the population parameters and
  the individual parameters.  This will then compare the results
  between `NONMEM` and `rxode2` to make sure the translation makes
  sense.  This only works when `nonmem2rx` has access to the input
  data and the output with the `IPRED`, `PRED` and the `ETA` values.

- Converts the upper case NONMEM variables to lower case.

- Replaces the NONMEM theta / eta names with the label-based names like an
  extended control stream

- Replaces the compartment names with the defined compartment names in
  the control stream (ie `COMP=(compartmenName)`)

### Comparing differences between `NONMEM` and `rxode2`

You may wish to see where the differences in predictions are between
NONMEM and rxode2.

The easiest way to visually compare the differences is by the plot method:

```{r plot}
plot(mod) # for general plot
# you can also see individual comparisons
plot(mod, log="y", ncol=2, nrow=2,
     xlab="Time (hr)", ylab="Concentrations",
     page=1)
```

If you want numerical differences, you can also get these from the
modified returned `ui` object.  For the rtol, atol as follows you have:

```{r tol}
mod$iwresAtol
mod$iwresRtol
mod$ipredAtol
mod$ipredRtol
mod$predAtol
mod$predAtol
```

You can see they do not exactly match but are very close (I would say
they validate). However you can explore these difference further if you wish
by looking at the `ipredCompare` and `predCompare` datasets:

```{r compare}
head(mod$iwresCompare)

head(mod$ipredCompare)

head(mod$predCompare)
```

In these cases you can see that NONMEM seems to round the values for
the output (the rounding rules are based on the `FORMAT` option), but
rxode2 seems to keep the entire number.

Note this is the **observation data only** that is compared.  Dosing
predictions are excluded from these comparisons.

You can also explore the NONMEM input dataset that was used to make
the validation predictions (dosing and observations) by the
`$nonmemData` item:

```{r nonmemData}
head(mod$nonmemData) # with nlme loaded you can also use getData(mod)
```

### Notes on validation

The validation of the model uses the best data available for NONMEM
estimates.  This is:

- `theta` or population parameters
- `eta` or individual parameters

The `omega` and `sigma` matrices are captured. When the nlmixr2 model
is fully qualified, the `IWRES` validation ensures the residual errors
are specified correctly. Otherwise `omega` and `sigma` values do not contribute to
the validation.  Also the overall covariance is captured, but not used
in the validation.

#### `theta` parameters:

This section discusses the source of the `theta` parameters.  Often it
is the same source of the `omega` and `sigma` matrices as well.  In
many cases it is the source of the overall covariance too.  So the
sources of all these data will be mentioned.

In order of preference, the model run parameters are gathered from:

- XML parameter estimates (which are the most accurate). On some
  systems, the `xml` may be broken (or not present).  This also gets
  the covariance, `omega`, and `sigma` values.

- `.ext` is the next source of the `theta`, `sigma` and `omega`
  matrices, and is slightly less accurate than the `xml` file.  If the
  covariance file `.cov` exists the overall covariance is read from
  this file when the `xml` file is not present.

- The NONMEM output file is the next source of information for the
  `theta` values.  This is the least accurate, but may be useful for
  models in libraries like the ddmore repository since often the
  `xml`, `ext` and `cov` files are not present.

- The model control stream is the last source of parameter estimates,
  and may not even reflect the minimum value of the model.  In common
  practice many people adjust parameters close to the final parameter
  estimates so this may be close enough, or a starting place to start
  your own estimation with `nlmxir2`.

### Eta parameters

The `eta` parameters come from 2 different sources.

- The first file (which is preferred) is used for classical estimation
  methods like `focei`.  In these circumstances the `.phi` file
  contains more accurate estimate information than the table files
  (without a `FORMAT` specifier).

- The next source of the `eta` parameters come from the tables.  The tables by
  default are not very accurate compared to the `xml` or `ext` files
  above.  The accuracy can be increased by changing the `FORMAT`
  option in the NONMEM tables.

Since these values are not as accurate as the population parameters,
these are much more likely to have a mismatch between `rxode2` and
`NONMEM`.  You can compare the predictions if you wish.

### What NONMEM and rxode2 compare

For the comparison we use the `PRED` and `IPRED` output in the tables
from NONMEM.

### Using `FORMAT` in nonmem

If you wish to be more precise for the `IPRED` and `ETA` values output
in your tables you can change the FORTRAN format to a something a bit
higher.

The default for tables is `FORMAT=s1PE11.4`, if you really want to
increase the precision for the output you could use `FORMAT=s1PE17.9`
or something similar, this can help with validation of individuals if
you get in a situation where it is important.

See [discussion on nmusers](mail-archive.com/nmusers@globomaxnm.com/msg05466.html).

### What do the values mean

Depending on the precision the validation numbers can be quite
different.  For example:

```{r validation}
# full parameter precision
mod <- suppressMessages(nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res"))

print(mod$ipredAtol)
print(mod$ipredRtol)

print(mod$predAtol)
print(mod$predRtol)

print(mod$iwresAtol)
print(mod$iwresRtol)

# now reduce precision by using table/lst output only
mod <- suppressMessages(nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res",
                                  useXml=FALSE, useExt=FALSE,usePhi=FALSE))

print(mod$predAtol)
print(mod$predRtol)

print(mod$ipredAtol)
print(mod$ipredRtol)

print(mod$iwresAtol)
print(mod$iwresRtol)
```

You can see that using less precise values will lead to larger
differences between the NONMEM model predictions and the rxode2 model
predictions. The key to validating models is to use the most precision
in parameter estimates (thetas and etas).

You can have situations where the model validates for the most precise
estimates, the population predictions, but doesn't seem to do a good
job for the individual estimates (which are a bit less precise since
they are not captured in the xml file). This could mean that one or more
`eta` values are highly sensitive to rounding.

Another possibility is that there are slight differences in how NONMEM
and rxode2 protects from division from zero and other possibilities that
can lead to `NaN` values.

If these model parameters are not as precise as you wish, try to get
more files, or maybe see if it still predicts the data accurately
enough by using the model to perform a `vpc`.

### Simulating from the model

Unlike the traditional `rxode2` `ui` objects, this contains
information about the model that can be used for simulation with
uncertainty: theta covariance, number of subjects and number of
observations.  You can see them below if you wish:

```{r simInfo}
mod$dfSub # Number of subjects
mod$dfObs # Number of observations
# Covariance of the thetas (and omegas in this case)
mod$thetaMat
```

# External contributors (and notes on validation)

This was validated against:

- The `PsN` library test suite of NONMEM listings
(https://github.com/UUPharmacometrics/PsN/tree/master/test)

- The ddmore model scrapings
  (https://github.com/dpastoor/ddmore_scraping).

- Models from NONMEM design tutorial Bauer 2021 https://doi.org/10.1002/psp4.12713

- Models from NONMEM tutorial 1 (Bauer 2019) https://doi.org/10.1002/psp4.12404

- Models from NONMEM tutorial 2 (Bauer 2019)  https://doi.org/10.1002/psp4.12422

Due to the sheer size of the zipped models for these nonmem control
stream sources, these are excluded to keep the binary below 3 mgs
(CRAN requirement).

However, I would like to acknowledge all who helped in these projects.
With these projects the NONMEM conversion to rxode2 has been made much
more robust.

Still, while the tests are not/will not be in the CRAN binaries, you
can test them yourself by:

1. Downloading this repository
2. Running the tests `devtools::test()`

