---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonmem2rx

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nonmem2rx/branch/main/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nonmem2rx?branch=main)
[![CRAN version](http://www.r-pkg.org/badges/version/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CodeFactor](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx/badge)](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx)
![r-universe](https://nlmixr2.r-universe.dev/badges/nonmem2rx)
<!-- badges: end -->

The goal of nonmem2rx is to convert a NONMEM control stream to a
`rxode2` (or even a `nlmixr2` fit) for easy clinical trial simulation
in R.

## Installation

You can install the development version of nonmem2rx from
[GitHub](https://github.com/) with the r-universe:

``` r
install.packages('nonmem2rx', repos = c('https://nlmixr2.r-universe.dev', 'https://cloud.r-project.org'))
```

When on CRAN, you can also get the CRAN version by:

``` r
install.packages('nonmem2rx')
```

## What you can do with `nonmem2rx`

## What you can do with `nonmem2rx`/`babelmixr2`

You can do many useful tasks directly converting between nlmixr2 and NONMEM models; you can:

-	[Convert a NONMEM model to a rxode2 model](https://nlmixr2.github.io/nonmem2rx/articles/import-nonmem.html)

-	[Do development in nlmixr2 and then run NONMEM from a nlmixr2 model](https://nlmixr2.github.io/babelmixr2/articles/running-nonmem.html) for reviewers who want to know about NONMEM results.

-	In both conversions, [automatically make sure the model is translated correctly ](https://nlmixr2.github.io/nonmem2rx/articles/rxode2-validate.html) (for [babelmixr2](https://nlmixr2.github.io/babelmixr2/articles/running-nonmem.html#optional-step-2-recover-a-failed-nonmem-run))

Then with nlmixr2 fit models coming from both conversions, you can:

-	[Perform simulations of new
    dosing](https://nlmixr2.github.io/nonmem2rx/articles/simulate-new-dosing.html) from the NONMEM model or even [simulate using the uncertainty in your model to simulate new scenarios](https://nlmixr2.github.io/nonmem2rx/articles/simulate-uncertainty.html)

-	[Modify the model to calculate derived parameters](https://nlmixr2.github.io/nonmem2rx/articles/simulate-extra-items.html) (like AUC). These parameters slow down NONMEM’s optimization, but can help in your simulation scenario.

With nonmem2rx and babelmixr2, convert the imported rxode2 model to a nlmixr2 object, allowing:

-	[Generation of Word and PowerPoint plots with
      nlmixr2rpt](https://nlmixr2.github.io/nonmem2rx/articles/create-office.html)

-	[Easy VPC
      creation](https://nlmixr2.github.io/nonmem2rx/articles/create-vpc.html)
      (with `vpcPlot()`)

-	[Easy Individual plots with extra solved
      points](https://nlmixr2.github.io/nonmem2rx/articles/create-augPred.html). This
      will show the curvature of individual and population fits for
      sparse data-sets (with `augPred()`)

You can even use this conversion to help debug your NONMEM model (or
even try it in nlmixr2 instead)

-	[Understand how to simplify the NONMEM model to avoid rounding errors](https://nlmixr2.github.io/nonmem2rx/articles/read-rounding.html)

-	[Run nlmixr2’s covariance step when NONMEMs covariance step has failed](https://nlmixr2.github.io/nonmem2rx/articles/read-rounding.html#step-5-get-the-covariance-of-the-model) (in the linked example, there was no covariance step because rounding errors)

    
## Simple example

Once `nonmem2rx` has been loaded, you simply type the location of the
nonmem control stream for the parser to start.  For example:

```{r example}
library(nonmem2rx)
mod <- nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res", save=FALSE)
mod
```

You can see this automatically validates NONMEM and rxode2 outputs for
a couple of metrics.

# External projects that contributed to the tool's validation

The `nonmem2rx` tool was validated against:

- The `PsN` library test suite of NONMEM listings
(https://github.com/UUPharmacometrics/PsN/tree/master/test)

- The ddmore model scrapings
  (https://github.com/dpastoor/ddmore_scraping).

- Models from NONMEM design tutorial Bauer 2021 https://doi.org/10.1002/psp4.12713

- Models from NONMEM tutorial 1 (Bauer 2019) https://doi.org/10.1002/psp4.12404

- Models from NONMEM tutorial 2 (Bauer 2019)  https://doi.org/10.1002/psp4.12422

Due to the sheer size of the zipped models for these nonmem control
stream sources, these are excluded to keep the binary below 3 mgs
(CRAN requirement).

However, I would like to acknowledge all who helped in these projects.
With these projects the NONMEM conversion to rxode2 has been made much
more robust.

Still, while the tests are not/will not be in the CRAN binaries, you
can test them yourself by:

1. Downloading this repository
2. Running the tests `devtools::test()`

