---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonmem2rx

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nonmem2rx/branch/main/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nonmem2rx?branch=main)
[![CRAN version](http://www.r-pkg.org/badges/version/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CodeFactor](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx/badge)](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx)
![r-universe](https://nlmixr2.r-universe.dev/badges/nonmem2rx)
<!-- badges: end -->

The goal of nonmem2rx is to convert a NONMEM control stream to
`rxode2` for easy clinical trial simulation in R.

## Installation

You can install the development version of nonmem2rx from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nlmixr2/nonmem2rx")
```

## Example

Once `nonmem2rx` has been loaded, you simply type the location of the
nonmem control stream for the parser to start.  For example:

```{r example}
library(nonmem2rx)
mod <- nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res")
mod 
```

The process steps are below:

- Read in the nonmem control stream and convert the model to a
  `rxode2` ui function.
  
- Try to determine an endpoint in the model (if possible), and convert
  to a fully qualified ui model that can be used in `nlmixr2` and
  `rxode2`.

- If available, `nonmem2rx` will read the final parameter estimates
  and update the model. (It tries the `.ext` file followed by the `.lst`
  file; without these files, will simply keep the initial estimates in
  the model).
  
- This will read in the nonmem input dataset, and search for the
  output files with `IPRED`, `PRED` and the `ETA` values. The
  translated `rxode2` model is run for the population parameters and
  the individual parameters.  This will then compare the results
  between `NONMEM` and `rxode2` to make sure the translation makes
  sense.  This only works when `nonmem2rx` has access to the input
  data and the output with the `IPRED`, `PRED` and the `ETA` values.
  
- Converts the upper case NONMEM variables to lower case.

- Replaces the NONMEM theta / eta names with the label-based names like an
  extended control stream
  
- Replaces the compartment names with the defined compartment names in
  the control stream (ie `COMP=(compartmenName)`)
  
### Comparing differences between `NONMEM` and `rxode2`

You may wish to see where the differences in predictions are between
NONMEM and rxode2.

From the modified returned `ui` object you can look at the rtol, atol as follows:

```{r tol}
mod$ipredAtol
mod$ipredRtol
mod$predAtol
mod$predAtol
```

You can see they do not exactly match. You can explore these
difference further if you wish by looking at the `ipredCompare` and
`predCompare` datasets:

```{r compare}
head(mod$ipredCompare)

head(mod$predCompare)
```

In these cases you can see that NONMEM seems to round the values for
the output (though I am not clear what the rounding rules are), but
rxode2 seems to keep the entire number.

Note this is the **observation data only** that is compared.  Dosing
predictions are excluded from these comparisons.

You can also explore the NONMEM input dataset that was used to make
the validation predictions (dosing and observations) by the
`$nonmemData` item:

```{r nonmemData}
head(mod$nonmemData) # with nlme loaded you can also use getData(mod)
```

### Notes on validation

The validation of the model uses the best data available for NONMEM
estimates.  This is:

- XML parameter estimates
- `.phi` estimates for etas (better than table outputs)

On some systems, the `xml` may be broken (or not present), then we use:

- `.ext`, and `.cov` parameter estimates
- `.phi` estimates for etas (better than table outputs)

If the `phi` isn't available, it will use the output from the tables
(if they include `ETAs`)

If none of these are available, then it will use the `.lst` file
estimates.

For the comparison we use the `PRED` and `IPRED` output in the table.

Depending on the precision the validation numbers can be quite
different.  For example:

```{r validation}
# full parameter precision
mod <- suppressMessages(nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res"))

print(mod$ipredAtol)
print(mod$ipredRtol)

print(mod$predAtol)
print(mod$predRtol)

# now reduce precision by using table/lst output only
mod <- suppressMessages(nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res",
                                  useXml=FALSE, useExt=FALSE,usePhi=FALSE))

print(mod$predAtol)
print(mod$predRtol)

print(mod$ipredAtol)
print(mod$ipredRtol)
```

You can see that using less precise values will lead to larger
differences between the NONMEM model predictions and the rxode2 model
predictions. The key to validating models is to use the most precision
possible. 

You can have situations where the model validates for the most precise
estimates, the population predictions, but doesn't seem to do a good
job for the individual estimates (which are a bit less precise since
they are not captured in the xml file). This could mean that one or more
`eta` values are highly sensitive to rounding.

Another possibility is that there are slight differences in how NONMEM
and rxode2 protects from division from zero and other possibilities that
can lead to `NaN` values.

If these model parameters are not as precise as you wish, try to get
more files, or maybe see if it still predicts the data accurately
enough by using the model to perform a `vpc`.

### Simulating from the model

Unlike the traditional `rxode2` `ui` objects, this contains
information about the model that can be used for simulation with
uncertainty: theta covariance, number of subjects and number of
observations.  You can see them below if you wish:

```{r simInfo}
mod$dfSub # Number of subjects
mod$dfObs # Number of observations
# Covariance of the thetas (and omegas in this case)
mod$thetaMat
```
