---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonmem2rx

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nonmem2rx/branch/main/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nonmem2rx?branch=main)
[![CRAN version](http://www.r-pkg.org/badges/version/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CodeFactor](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx/badge)](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx)
![r-universe](https://nlmixr2.r-universe.dev/badges/nonmem2rx)
<!-- badges: end -->

The goal of nonmem2rx is to convert a NONMEM control stream to
`rxode2` for easy clinical trial simulation in R.

## Installation

You can install the development version of nonmem2rx from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nlmixr2/nonmem2rx")
```

## Example

Once `nonmem2rx` has been loaded, you simply type the location of the
nonmem control stream for the parser to start.  For example:

```{r example}
library(nonmem2rx)
mod <- nonmem2rx(system.file("run001.mod", package="nonmem2rx"))
mod 
```

The process steps are below:

- Read in the nonmem control stream and convert the model to a
  `rxode2` ui function.
  
- Try to determine an endpoint in the model (if possible), and convert
  to a fully qualified ui model that can be used in `nlmixr2` and
  `rxode2`.

- If available, `nonmem2rx` will read the final parameter estimates
  and update the model. (It tries the `.ext` file followed by the `.lst`
  file; without these files, will simply keep the initial estimates in
  the model).
  
- This will read in the nonmem input dataset, and search for the
  output files with `IPRED`, `PRED` and the `ETA` values. The
  translated `rxode2` model is run for the population parameters and
  the individual parameters.  This will then compare the results
  between `NONMEM` and `rxode2` to make sure the translation makes
  sense.  This only works when `nonmem2rx` has access to the input
  data and the output with the `IPRED`, `PRED` and the `ETA` values.
  
- Converts the upper case NONMEM variables to lower case.

- Replaces the NONMEM theta / eta names with the label-based names like an
  extended control stream
  
- Replaces the compartment names with the defined compartment names in
  the control stream (ie `COMP=(compartmenName)`)

