---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonmem2rx

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nonmem2rx/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nonmem2rx/branch/main/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nonmem2rx?branch=main)
[![CRAN version](http://www.r-pkg.org/badges/version/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/nonmem2rx)](https://cran.r-project.org/package=nonmem2rx)
[![CodeFactor](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx/badge)](https://www.codefactor.io/repository/github/nlmixr2/nonmem2rx)
![r-universe](https://nlmixr2.r-universe.dev/badges/nonmem2rx)
<!-- badges: end -->

The goal of nonmem2rx is to convert a NONMEM control stream to
`rxode2` for easy clinical trial simulation in R.

## Installation

You can install the development version of nonmem2rx from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nlmixr2/nonmem2rx")
```

## Example

Once `nonmem2rx` has been loaded, you simply type the location of the
nonmem control stream for the parser to start.  For example:

```{r example}
library(nonmem2rx)
mod <- nonmem2rx(system.file("mods/cpt/runODE032.ctl", package="nonmem2rx"), lst=".res")
mod 
```

The process steps are below:

- Read in the nonmem control stream and convert the model to a
  `rxode2` ui function.
  
- Try to determine an endpoint in the model (if possible), and convert
  to a fully qualified ui model that can be used in `nlmixr2` and
  `rxode2`.

- If available, `nonmem2rx` will read the final parameter estimates
  and update the model. (It tries the `.ext` file followed by the `.lst`
  file; without these files, will simply keep the initial estimates in
  the model).
  
- This will read in the nonmem input dataset, and search for the
  output files with `IPRED`, `PRED` and the `ETA` values. The
  translated `rxode2` model is run for the population parameters and
  the individual parameters.  This will then compare the results
  between `NONMEM` and `rxode2` to make sure the translation makes
  sense.  This only works when `nonmem2rx` has access to the input
  data and the output with the `IPRED`, `PRED` and the `ETA` values.
  
- Converts the upper case NONMEM variables to lower case.

- Replaces the NONMEM theta / eta names with the label-based names like an
  extended control stream
  
- Replaces the compartment names with the defined compartment names in
  the control stream (ie `COMP=(compartmenName)`)
  
### Comparing differences between `NONMEM` and `rxode2`

You may wish to see where the differences in predictions are between
NONMEM and rxode2.

From the modified returned `ui` object you can look at the rtol, atol as follows:

```{r tol}
mod$ipredAtol
mod$ipredRtol
mod$predAtol
mod$predAtol
```

You can see they do not exactly match. You can explore these
difference further if you wish by looking at the `ipredCompare` and
`predCompare` datasets:

```{r compare}
head(mod$ipredCompare)

head(mod$predCompare)
```

In these cases you can see that NONMEM seems to round the values for
the output (though I am not clear what the rounding rules are), but
rxode2 seems to keep the entire number.

Note this is the **observation data only** that is compared.  Dosing
predictions are excluded from these comparisons.

You can also explore the NONMEM input dataset that was used to make
the validation predictions (dosing and observations) by the
`$nonmemData` item:

```{r nonmemData}
head(mod$nonmemData) # with nlme loaded you can also use getData(mod)
```

### Simulating from the model

Unlike the traditional `rxode2` `ui` objects, this contains
information about the model that can be used for simulation with
uncertainty: theta covariance, number of subjects and number of
observations.  You can see them below if you wish:

```{r simInfo}
mod$dfSub # Number of subjects
mod$dfObs # Number of observations
# Covariance of the thetas (and omegas in this case)
mod$thetaMat
```
