xxxxx---
title: "Simulate New dosing with covariates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulate New dosing with covariates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Simulation with covariates or input parameters

Sometimes your NONMEM model can have covariates that you may wish to
simulate from; this simulation exercise shows a few methods to
simulate with the covariates from NONMEM.

```{r setup}
library(nonmem2rx)
library(rxode2)
```

## Step 0: input the model

In this case, we will use the Friberg myelosuppresion model originally
contributed as an example by Yuan Xiong.

With the simulated data in nlmixr2, babelmixr2, and some manual edits
to simplify the model we run NONMEM 7.4.3.  

Note in this case there are some PK parameters that are in the model
that require some special handling to simulate with uncertainty or
even with different dosing scenarios.

For any simulation scenario, we need to import the NONMEM model:

```{r wbcImport}
# Since this is an included example, we import the model from the
# `nonmem2rx` package.  This is done by the `system.file()` command:
wbcModel <- system.file("wbc/wbc.lst", package="nonmem2rx")

wbc <- nonmem2rx(wbcModel)

print(wbc)

# note the NONMEM vs rxode2 models validate well. You can see this in
# the validation code:
message(paste(wbc$meta$validation, collapse="\n"))
```

## Option #1: simulate with the same conditions as the input model

The easiest way to simulate with uncertainty is to use the original
NONMEM input dataset.  If we want to simulate covariates from here, we
simply add `resample=TRUE`:

```{r simCovUncertainty}
sim <- rxSolve(wbc, resample=TRUE, nStud=500)
```

In this case every individual re-samples from the original dataset's
covariates.  In this particular case, the dosing changes per
individual and it may not be what you wish to share with the team but
may be a way to see how the model is performing relative to the data.
Binning may be necessary, as with a typical `VPC`

## Option 2: simulate with a different condition (with resampled PK parameters/covariates)

In this case, you may wish to simulate a study that has similar
covariates as the NONMEM model in general (and also with resampling)

First lets simulate `410` every 20 days.  We can easily add this by
creating a event table with the same input PK parameters as the NONMEM
dataset.


```{r simCovUncertainty2}
# first create the base event table with the nubmer of individuals
# matching the NONMEM dataset:
ev <- et(amt=410, ii=20*24, until=365*24) %>% # Add dosing 20 days apart for a year
  et(seq(0, 365*24, by=7*24)) %>% # Assume weekly observations
  et(id=seq_along(unique(wbc$nonmemData$ID))) %>% # Match the number of subjects modeled
  as.data.frame # convert to data.frame

# Now create the PK covariates
library(dplyr)
pkCov <- wbc$nonmemData %>%
  filter(!duplicated(ID)) %>% # only get one observation per id
  select(CLI, V1I, V2I) # select the covariates

pkCov$id <- seq_along(pkCov$CLI) # add the covariates per id

# Then merge the PK covariates to the original event table
ev <- merge(pkCov, ev)

# Last simulate with replacement with the new data frame

sim <- rxSolve(wbc, ev, resample=TRUE, nStud=100)

ci <- confint(sim, "y")

plot(ci)
```

This may be closer a constant theoretical dosing regimen you may wish
to explore. 

## Option 3: simulate a larger study with a different condition (resampled PK parameters/covariates)

Another option is to create a larger dataset (that is a multiple of
the original dataset). In this case, I will assume that the new study
will have `225` patients, which is a `5` fold increase in subjects
compared to the NONMEM input.

```{r simUncertanty3}
# first create the base event table with the nubmer of individuals
# matching the NONMEM dataset:
ev <- et(amt=410, ii=20*24, until=365*24) %>% # Add dosing 20 days apart for a year
  et(seq(0, 365*24, by=7*24)) %>% # Assume weekly observations
  et(id=seq(1, max(wbc$nonmemData$ID)*5)) %>% # Match the number of subjects modeled
  as.data.frame # convert to data.frame

# Now create the PK covariates
library(dplyr)

pkCov <- wbc$nonmemData %>%
  filter(!duplicated(ID)) %>% # only get one observation per id
  select(CLI, V1I, V2I) # select the covariates

# expand the covariates by 5

pkCov <- do.call("rbind",
                 lapply(1:5, function(i) {
                   pkCov
                 }))


pkCov$id <- seq_along(pkCov$CLI) # add the covariates per id

# Then merge the PK covariates to the original event table
ev <- merge(pkCov, ev)

# Last simulate with replacement with the new data frame

sim <- rxSolve(wbc, ev, resample=TRUE, nStud=100)

ci <- confint(sim, "y")

plot(ci)
```

## Other options

You can also simulation without uncertainty and use covariates by
resampling by hand (or even simulating new covariates manually).

I believe that reampling keeps any hidden correlations between
covariates, and should be used whenever possible.  At the time of this
writing, resampling can only occur when the new event table is a
multiple of the input dataest.  Eventually a feature may be added to
resample from an input dataset directly.

Note that resampling will also work with time-varying covariates.  The
time-varying covariates would be imputed based on the input times per
subject.
