---
title: "Simulate using Parameter Uncertainty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulate using Parameter Uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This page shows a simple work-flow for directly simulating a different
dosing paradigm than what was modeled taking into account the modeled
uncertainty. This workflow is very similar to simply [simulating
without uncertainty](simulate-new-dosing.html) in the parameters
themselves.

## Step 1: Import the model

```{r setup}
library(nonmem2rx)
library(rxode2)
# its best practice to set the seed for the simulations
set.seed(42)
rxSetSeed(42)

# First we need the location of the nonmem control stream Since we are
# running an example, we will use one of the built-in examples in
# `nonmem2rx`
ctlFile <- system.file("mods/cpt/runODE032.ctl", package="nonmem2rx")
# You can use a control stream or other file. With the development
# version of `babelmixr2`, you can simply point to the listing file
mod <- nonmem2rx(ctlFile, lst=".res", save=FALSE, determineError=FALSE)
```

## Step 2: Look at a different dosing paradigm

Lets say that in this case instead of a single dose, we want to see
what the concentration profile is with a single day of BID dosing.  In
this case is done by creating a [quick event
table](https://nlmixr2.github.io/rxode2/articles/rxode2-event-table.html).

```{r eventTable}
ev <- et(amt=120000, ii=12, until=24) %>%
  et(c(1:6, seq(8, 24, by=2))) %>%
  et(id=1:100)
```


## Step 3: Solve using the uncertainty in the NONMEM model

To use the uncertainty in the model, it is a simple matter of telling
how many times `rxode2()` should sample with `nStud=X`.  In this case
we will use `100`.

```{r rxSolve}
s <- rxSolve(mod, ev, nStud=100)

s
```

## Step 4: Summarize and plot

Since there is a bunch of data, a confidence band of the simulation with uncertainty would be helpful. 

One way to do that is to select the interesting components, create a confidence interval and then plot the confidence bands:



```{r confint}
sci <- confint(s, parm=c("CENTRAL", "PERI", "sim"))

sci

plot(sci)

plot(sci, log="y")
```
