---
title: "Converting a NONMEM fit to a nlmixr2 object"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### Creating a nlmixr2 compatible model 

Depending on the model, not all the residual specifications are
translated to the `nlmixr2` style residuals.  This means the model
cannot be immediately used for either `nlmixr2()` estimation or
creating a `nlmixr2` fit object (though you can simulate with and
without certainty without any modifications)

For example you could have something like:

```r
y <- ipred*(1+eps1)
```

For a model that can do `nlmixr2` estimation instead of simply
simulation the residual needs to be changed to something like:

```r
cp ~ prop(prop.sd)
```

Since the model when import has most of the translation done already,
you can easily tweak the model to have this form. 

Here is the same example where the residual errors are not
automatically translated to the `nlmixr2` parameter style (in this
case because of the option `determineError=FALSE`)

## Example -- no error determined

```{r asNonmem2Rx}
library(nonmem2rx)
library(babelmixr2)

# First we need the location of the nonmem listing file
# Since we are running an example, we will use one of the built-in examples in `nonmem2rx`
lstFile <- system.file("mods/cpt/runODE032.res", package="nonmem2rx")
# You can use a control stream or other file, and change the `lst`
# import too, but this seems like the easiest to me.

# Of course you woud use your own listing file
mod <- nonmem2rx(lstFile, save=FALSE, determineError=FALSE)

print(mod)
```

One approach to get a nlmixr2 compatible model is to copy the printed
model above and then modify it as needed.

In this case, I will name the parameters to something a bit more
meaningful while keeping the estimates the same:

```{r manuallySpecifyModel}
mod2 <-function() {
  ini({
    lcl <- 1.37034036528946
    lvc <- 4.19814911033061
    lq <- 1.38003493562413
    lvp <- 3.87657341967489
    RSV <- c(0, 0.196446108190896, 1)
    eta.cl ~ 0.101251418415006
    eta.v ~ 0.0993872449483344
    eta.q ~ 0.101302674763154
    eta.v2 ~ 0.0730497519364148
  })
  model({
    cmt(CENTRAL)
    cmt(PERI)
    cl <- exp(lcl + eta.cl)
    v <- exp(lvc + eta.v)
    q <- exp(lq + eta.q)
    v2 <- exp(lvp + eta.v2)
    v1 <- v
    scale1 <- v
    k21 <- q/v2
    k12 <- q/v
    d/dt(CENTRAL) <- k21 * PERI - k12 * CENTRAL - cl * CENTRAL/v1
    d/dt(PERI) <- -k21 * PERI + k12 * CENTRAL
    f <- CENTRAL/scale1
    f ~ prop(RSV)
  })
}
```

The `as.nonmem2rx()` function will compare the already imported rxode2
model function with the model you made some manual tweaks to:

```{r}
new <- as.nonmem2rx(mod2, mod)

print(new)
```

In this case the `new` model qualifies and now has all the information
from the imported nonmem2rx model.

This means you can estimate from the new model knowing it was the same
model specified in NONMEM.

Since `iwres` is affected by how your specify your residuals, pay
special attention to that validation.  If it does not validate, you
may have forgot to translate the NONMEM variance estimate to the
standard deviation estimate required by many estimation methods.


### Converting the model to a nlmixr2 fit

Once you have a `rxode2()` model that:

- Qualifies against the NONMEM model, 

- Has `nlmixr2` compatible residuals

You can then convert it to a `nlmixr2` fit object with `babelmixr2`:

```{r convertNlmixr2object}
library(babelmixr2)
fit <- as.nlmixr2(new)

fit
```
