---
title: "Create PowerPoint and Word documents using nonmem2rx"
resource_files:
  - mod-PowerPoint.pptx
  - mod-Word.docx
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Step 1: import the model into `nonmem2rx`

```{r setup}
library(nonmem2rx)
library(babelmixr2)
library(nlmixr2rpt)
library(onbrand)

library(nonmem2rx)

# First we need the location of the nonmem control stream Since we are running an example, we will use one of the built-in examples in `nonmem2rx`
ctlFile <- system.file("mods/cpt/runODE032.ctl", package="nonmem2rx")
# You can use a control stream or other file. With the development
# version of `babelmixr2`, you can simply point to the listing file

mod <- nonmem2rx(ctlFile, lst=".res", save=FALSE)
```


# Step 2: convert the `rxode2` model to `nlmixr2`

In this step, you convert the model to `nlmixr2` by `as.nlmixr2(mod)`;
You may need to do a [little manual work to get the residual
specification to match between nlmixr2 and NONMEM](convert-nlmixr2.html).

Once the residual specification is compatible with a nlmixr2 object, you can convert the model, `mod`, to
a nlmixr2 fit object:

```{r toNlmixr2}
fit <- as.nlmixr2(mod)

fit
```

# Step 3: Create a PowerPoint file

A PowerPoint can be created from your own [custom powerpoint
templates](https://nlmixr2.github.io/nlmixr2rpt/articles/Reporting_nlmixr_Fit_Results.html#customizing-reports-for-your-organization),
but in this example we will use the ones that come from `nlmixr2rpt`
directly:

```{r toPptx}
obnd_pptx = read_template(
  template = system.file(package="nlmixr2rpt", "templates","nlmixr_obnd_template.pptx"),
  mapping  = system.file(package="nlmixr2rpt", "templates","nlmixr_obnd_template.yaml"))

obnd_pptx = report_fit(
  fit     = fit, 
  obnd    = obnd_pptx)

save_report(obnd_pptx, "mod-PowerPoint.pptx")
```

Which gives the powerpoint [here](mod-PowerPoint.pptx)


# Step 4: Create a Word file

Just like in PowerPoint, you can customizeown [custom word templates](https://nlmixr2.github.io/nlmixr2rpt/articles/Reporting_nlmixr_Fit_Results.html#customizing-reports-for-your-organization), but in this example we will use the ones that come from `nlmixr2rpt` directly:


```{r}
obnd_docx = read_template(
  template = system.file(package="nlmixr2rpt", "templates","nlmixr_obnd_template.docx"),
  mapping  = system.file(package="nlmixr2rpt", "templates","nlmixr_obnd_template.yaml"))

obnd_docx = report_fit(
  fit     = fit, 
  obnd    = obnd_docx)

save_report(obnd_docx, "mod-Word.docx")
```

Which gives the word document [here](mod-Word.docx)
