---
title: "Easily Create a VPC using nonmem2rx"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
This shows an easy work-flow to create a VPC using a NONMEM model:

# Step 1: Convert the `NONMEM` model to `rxode2`:

```{r setup}
library(babelmixr2)
library(nonmem2rx)



# First we need the location of the nonmem control stream Since we are running an example, we will use one of the built-in examples in `nonmem2rx`
ctlFile <- system.file("mods/cpt/runODE032.ctl", package="nonmem2rx")
# You can use a control stream or other file. With the development
# version of `babelmixr2`, you can simply point to the listing file

mod <- nonmem2rx(ctlFile, lst=".res", save=FALSE)
```


# Step 2: convert the `rxode2` model to `nlmixr2`

In this step, you convert the model to `nlmixr2` by `as.nlmixr2(mod)`;
You may need to do a [little manual work to get the residual
specification to match between nlmixr2 and NONMEM](convert-nlmixr2.html).

Once the residual specification is compatible with a nlmixr2 object, you can convert the model, `mod`, to
a nlmixr2 fit object:

```{r toNlmixr2}
fit <- as.nlmixr2(mod)

fit
```

# Step 3: Perform the VPC

From here we simply use `vpcPlot()` in conjunction with the `vpc`
package to get the regular and prediction-corrected VPCs and arrange
them on a single plot:

```{r vpc}

library(ggplot2)
p1 <- vpcPlot(fit, show=list(obs_dv=TRUE))

p1 <- p1 + ylab("Concentrations") +
  rxode2::rxTheme() +
  xlab("Time (hr)") +
  xgxr::xgx_scale_x_time_units("hour", "hour")

p1a <- p1 + xgxr::xgx_scale_y_log10()

## A prediction-corrected VPC
p2 <- vpcPlot(fit, pred_corr = TRUE, show=list(obs_dv=TRUE))
p2 <- p2 + ylab("Prediction-Corrected Concentrations") +
  rxode2::rxTheme() +
  xlab("Time (hr)") +
  xgxr::xgx_scale_x_time_units("hour", "hour")

p2a <- p2 + xgxr::xgx_scale_y_log10()


library(patchwork)
(p1 * p1a) / (p2 * p2a)
```
