---
title: "Reading rounding from NONMEM"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

One of the things that can be helpful is reading NONMEM with rounding
errors.

For this example we will take some of the test examples from
`babelmixr2` to translate to show examples of rounding errors that can
be read in to help diagnose what could be happening with the model.

```{r setup}
library(nonmem2rx)
library(babelmixr2)
```

# Step 1: Have a NONMEM model with rounding errors (and `.phi` or other information about the `etas`)

The first step is to load a model with rounding errors using `nonmem2rx()`:

```{r loadNonmemModel}
# Unzip example with rounding error
# included, but can be accessed with nlmixr2
#
# unzip(system.file("tests/testthat/pk.turnover.emax3-nonmem.zip", package="babelmixr2"))
# Load the model with `nonmem2rx`:
mod <- nonmem2rx("pk.turnover.emax3-nonmem/pk.turnover.emax3.nmctl")
```

# Step 2: Convert `rxode2` model to model with endpoints/residuals specified like `nlmixr2`

This example has rounding errors and isn't a fully qualified `nlmixr2`
model (even though it was generated from `nlmixr2`).

We can use the model and create an equivalent model with
`as.nonmem2rx()`

```{r createNlmixr2Model}
mod2 <- function() {
  ini({
    tktr <- 6.24053043162953e-07
    label("1 - tktr")
    tka <- -3.00642760553675e-06
    label("2 - tka")
    tcl <- -2.00405074386117
    label("3 - tcl")
    tv <- 2.05188410700476
    label("4 - tv")
    prop.err <- c(0, 0.0985804613565218)
    label("5 - prop.err")
    pkadd.err <- c(0, 0.511625249037084)
    label("6 - pkadd.err")
    temax <- 6.4184983102259
    label("7 - temax")
    tec50 <- 0.140763261319656
    label("8 - tec50")
    tkout <- -2.9534704318737
    label("9 - tkout")
    te0 <- 4.57045413136592
    label("10 - te0")
    pdadd.err <- c(0, 3.71714384851537)
    label("11 - pdadd.err")
    eta.ktr ~ 0.558129815059436
    eta.ka ~ 0.558402321309217
    eta.cl ~ 0.0785849119252598
    eta.v ~ 0.0508226905750953
    eta.emax ~ 5e-05
    eta.ec50 ~ 0.18426809257979
    eta.kout ~ 0.0083631531443303
    eta.e0 ~ 0.00274561514766752
  })
  model({
    cmt(DEPOT)
    cmt(GUT)
    cmt(CENTER)
    cmt(EFFECT)
    ktr <- exp(tktr + eta.ktr)
    ka <- exp(tka   + eta.ka)
    cl <- exp(tcl   + eta.cl)
    v <- exp(tv     + eta.v)
    emax <- expit(temax + eta.emax) 
    ec50 <- exp(tec50 + eta.ec50)
    kout <- exp(tkout + eta.kout)
    e0 <- exp(te0 + eta.e0)
    EFFECT(0) <- e0
    dcp <- CENTER/v
    pd <- 1 - emax * dcp/(ec50 + dcp)
    kin <- e0 * kout
    d/dt(DEPOT) <- -ktr * DEPOT
    d/dt(GUT) <- ktr * DEPOT - ka * GUT
    d/dt(CENTER) <- ka * GUT - cl/v * CENTER
    d/dt(EFFECT) <- kin * pd - kout * EFFECT
    eff <- EFFECT
    dcp ~ add(pkadd.err)+prop(prop.err)
    eff ~ add(pdadd.err)
  })
}
```

In the model above, the following modifications were made:

- The code for protecting for rounding errors was removed

- Endpoints/residual specifications were added

- Duplicate code in NONMEM's `$ERROR` block was removed from imported
  model
  
- All the `rx` and `nlmixr` prefixed items were removed from the
  model.  It is possible that not removing these variables can
  cause`nlmixr2` conversion to fail.  It is best practice to remove
  them completely.

Also, it is good practice to make sure the model parses correctly before
trying to validate/convert the model.  If you find that your model doesn't
parse correctly it definitely won't validate (and the error may not be
as easy to track-down).

The first step is to validate if the translation is correct.  This is
done below:

```{r convertAsNonmem2rx}
mod2 <- mod2()

mod2

new <- as.nonmem2rx(mod, mod2)
print(new)
```

# Step 3: Convert new model to nlmixr2 fit with `as.nlmixr2()`

Once the translation is complete, and it is validated, you can convert
the fit to a full nlmixr2 fit object.

```{r convertAsNlmixr2}
fit <- as.nlmixr2(new)

# Once it is loaded remove the directory (we don't need the files any
# more for this example)
#
# In this example, we don't remove, but not where it can be removed
#
# unlink("pk.turnover.emax3-nonmem", recursive = TRUE)
```

Note this object **does not** rerun an estimation, but rather imports
everything it knows about

# Step 4: Explore the data/fit as if it came from `nlmixr2`

You can get the covariance with `getVarCov()` see:

```{r diagnoseProblems}
print(fit)

getVarCov(fit)
```

Then you can see the standard error here:

``{r standardError}
fit
```
